### 2014-11-20 src/1_Introduction.tex
avance

\section{Scope of this Research}

@@

This idea was inmediatly adopted in finance since it could represent their
long-run relationship implied by economic theory~\cite{laietAl1991},
\cite{lence+falk2005}.  Economic theory suggest that economic time series are
mean-reverting process and therefore, it reflects the idea of that some set of
variables cannot wander too far from each other.

On the other hand, the efficient markets hypothesis, also known as the random
walk theory states that current stock prices fully reflect available
information related to its value and there is no way to earn excess
profits~\cite{fama1970}.  This means that if we have stock prices from a
jointly efficient market, they cannot be
cointegrated~\cite{granger1986},\cite{dwyer1992}. However, \cite{richards1995}
claims that cointegration is directly at odds with market efficiency, even
though, there is no evidende that cointegration among asset prices have
implications about market efficiency~\cite{lence+falk2005}.

Despite the fact that cointegration on closing daily rates of currency pairs has
not been found \cite{coleman1990}, \cite{copeland1991}, different time series
frequencies can have different behaviours~\cite{aldridge2009}. Pair trading is a
very common example of cointegration application~\cite{herlemont2003} 
but cointegration can also be extended to a larger set of
variables~\cite{mukherjee1995},\cite{engle2004}.

%So why do we care about cointegration? Someone else can probably give more
%econometric applications, but in quantitative finance, cointegration forms the
%basis of the pairs trading strategy: suppose we have two cointegrated stocks X
%and Y, with the particular (for concreteness) cointegrating relationship X - 2Y
%= Z, where Z is a stationary series of zero mean. For example, X could be
%McDonald's, Y could be Burger King, and the cointegration relationship would
%mean that X tends to be priced twice as high as Y, so that when X is more than
%twice the price of Y, we expect X to move down or Y to move up in the near
%future (and analogously, if X is less than twice the price of Y, we expect X to
%move up or Y to move down). This suggests the following trading strategy: if X -
%2Y > d, for some positive threshold d, then we should sell X and buy Y (since we
%expect X to decrease in price and Y to increase), and similarly, if X - 2Y < -d,
%then we should buy X and sell Y.

@@

\section{Research Objectives}
The main motivation for this research is ...

The specific objectives of this research are,
\begin{itemize}
\item $\Large \mathcal{O}_1$: \emph{Especific Objective 1.}
\item $\Large \mathcal{O}_2$: \emph{Especific Objective 2.}
\item $\Large \mathcal{O}_3$: \emph{Especific Objective 3.}
\item $\Large \mathcal{O}_4$: \emph{Especific Objective 4.}
\item $\Large \mathcal{O}_5$: \emph{Especific Objective 5.}
\end{itemize}

The main research hypothesis explored in this dissertation is the following \\

\textit{Hypothesis llll . }

@@

Chapter 1 contains ...

>>>>

\section{Scope of this research}

The stochastic behaviour of financial time series, its incrementing
amount of data available and the need of performing accurate forecasting in
short periods of time has motivated researchers to create efficient and fast
algorithms. This study involves interdisciplinary knowledge such as: finance,
scientific computing, high performance computing, machine learning, physics
among others.

Forecasting financial time series have been modeled using classical statistical
models. More recently , machine learning models have been extensively used in
forecasting. However, getting model parameters is a computational challenge.
The computational complexity of machine learning algorithms has become a
limiting factor for problems that require processing large volumes of data and
where the response time is crucial.

Therefore, algorithms that process large amount of data in a short period of
time are required.  Online learning algorithms have been developed to solve
large-scale problems since they process an instance at a time, updating the
model at each step incrementally. This is opposed to the batch algorithms where
a phase of training that uses a large collection of historical data is first
required and the resulting model is later used to predict.

The specific scope of this study is to use financial time series features in
order to design a forecasting algorithm which ensures accuracy and low response
times. Cointegration is the main feature studied and it refers that one or more
linear combinations of these variables are stationary even though individually
they are not~\cite{engle87}.  Some models, such as the Vector Error Correction
(VECM), take advantage of this property and describe the joint behaviour of
several cointegrated variables.

\section{Research Objectives}
The main motivation for this research is the develpment of efficient methods for
forecasting financial time series.

The specific objectives of this research are,
\begin{itemize}
\item $\Large \mathcal{O}_1$: \emph{A review of the literature on time series
analysis models including machine learning techniques.}
\item $\Large \mathcal{O}_2$: \emph{Development of a set of know features of the
studied time series and applied them to improve forecasting.}
\item $\Large \mathcal{O}_3$: \emph{Development of parallel and efficient
algorithms to ensure quick time responses.}
\item $\Large \mathcal{O}_4$: \emph{Deep mathematical analysis of the proposal
and financial concepts involved.}
\item $\Large \mathcal{O}_5$: \emph{Design and implement representative set of
experiments in order to show when and why proposal performs better.}
\end{itemize}

@@

@@

The main research hypothesis explored in this dissertation is the following:
\\

\textit{An online learning algorithms based on cointegration and high
performance computing will allow to obtain better and faster forecasting
algorithms for financial time series maintaining good accuracy levels.}

@@

Chapter 1 contains relevant finance concepts required to understand financial
time series models such as market hypothesis, frequency, order book generation,
market microestructure.
Chapter 2 discuss main time series models including classic ones such as ARMA,
GARCH and more recent ones such as VECM, VAR and volatility models. %mejorar
Chapter 3 gives a brief introduction to machine learning and its variant online
learning. 
Chapter 4 presents different algorithms propose to forecast financial time
series using machine learning techniques.
In Chapter 5 a deep study of foreign exchange market features is presented and
experiments of cointegration are carried out in order to show its dynamic.
Chapter 6 presents a discussion of found results and summarize the main
conlusions of this thesis. It also presents some research directions for future
studied.

### 2014-11-20 src/3_TimeSeries.tex
avance

\chapter{Time Series}
\section{Implied volatility models}

>>>>

\chapter{Time Series}

Cointegration concept was introduced by~\cite{engle1987} and implies that one or
more linear combinations of non-stationary variables are stationary even though
individually they are not.  Moreover ~\cite{stock+watson1988} observed that
cointegration reflects the common stochastic trends providing a useful way to
understand cointegration relationships. These common stochastic trends can be
also interpreted as a long-run equilibrium relationships.

This idea was inmediatly adopted in finance since it could represent their
long-run relationship implied by economic theory~\cite{laietAl1991},
\cite{lence+falk2005}.  Economic theory suggest that economic time series are
mean-reverting process and therefore, it reflects the idea of that some set of
variables cannot wander too far from each other.

On the other hand, the efficient markets hypothesis, also known as the random
walk theory states that current stock prices fully reflect available
information related to its value and there is no way to earn excess
profits~\cite{fama1970}.  This means that if we have stock prices from a
jointly efficient market, they cannot be
cointegrated~\cite{granger1986},\cite{dwyer1992}. However, \cite{richards1995}
claims that cointegration is directly at odds with market efficiency, even
though, there is no evidende that cointegration among asset prices have
implications about market efficiency~\cite{lence+falk2005}.

Despite the fact that cointegration on closing daily rates of currency pairs has
not been found \cite{coleman1990}, \cite{copeland1991}, different time series
frequencies can have different behaviours~\cite{aldridge2009}. Pair trading is a
very common example of cointegration application~\cite{herlemont2003} 
but cointegration can also be extended to a larger set of
variables~\cite{mukherjee1995},\cite{engle2004}.
\chapter{Time Series}
\section{Implied volatility models}
